---
title: "final markdown"
author: "Ozan"
date: "2023-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BIGQA - CM Markdown file

## Importing libraries and functions from source

```{r}
library(Seurat)
library(tidyverse)
source("https://raw.githubusercontent.com/yasinkaymaz/BIGQA-CM/main/BIGQA-WP1-functions.R")
```

## Loading data

Original data contains whole datasets and cells and there is a subset of whole dataset thats volume is 2K cells.

**SOW-1** <https://drive.google.com/drive/folders/1yvK1RHT4fut96rFHQkHdrQGmqsxvThrw>

```{r}
#Load original data 
#seuObj <- readRDS("C:/Users/OZAN//Downloads/Bachireddy.all.tcells.seurat.rds")
#Load 2K cells subsample data:
seu <- readRDS("C:/Users/OZAN/Downloads/downsampled.2K.Bachireddy.all.tcells.seurat.rds")
```

## Alternative ways

#### 1. BISCUIT dataset loading process for 1K cell

```{r}
bis_data <- read.delim("~/Downloads/1000Tcells.biscuitNormalized.gem.txt",header = TRUE, sep = "\t", row.names = 1)
seu <- CreateSeuratObject(counts=bis_data, project="bisc", min.cells = 3, min.features = 200)
seu_info<- data.frame(read.delim("~/Downloads/1000Tcells_labels.txt", header = TRUE, sep = "\t"))
View(seu_info)
rownames(seu_info) <- seu_info[,1]
seu_info <- seu_info[,2:1000]
```

#### 2. Seurat processing

**Download raw count matrix**

<https://drive.google.com/drive/folders/1vMHAPbBSRouoyRymJ0cFdH1OBooYmK9f> **Note:** Cells were down-sampled to 1000 to recapitulate the method in the Bachireddy et al.

```{r}
cnt_b1 <- read.csv("/Users/yasinkaymaz/Documents/BIGQAWP1/cmlnf_B1_dense.csv", header=T, row.names = 1)
cnt_b1 <- t(cnt_b1[sample(nrow(cnt_b1), 1000), ])
seub1 <- CreateSeuratObject(counts = cnt_b1, project = "B01", min.cells = 3, min.features = 200)
seub1@meta.data$Outcome <- c("Responder")
seub1@meta.data$Pid <- c("5310")
seub1@meta.data$sampid <- c("B01")
seub1@meta.data$time <- c("pre")
rm(cnt_b1)
gc()
cnt_b3 <- read.csv("/Users/yasinkaymaz/Documents/BIGQAWP1/cmlnf_B3_dense.csv", header=T, row.names = 1)
cnt_b3 <- t(cnt_b3[sample(nrow(cnt_b3), 1000), ])
seub3 <- CreateSeuratObject(counts = cnt_b3, project = "B03", min.cells = 3, min.features = 200)
seub3@meta.data$Outcome <- c("Non-Responder")
seub3@meta.data$Pid <- c("5322")
seub3@meta.data$sampid <- c("B03")
seub3@meta.data$time <- c("pre")
head(seub3@meta.data)
rm(cnt_b3)
gc()
#Standard data processing by Seurat
seu <- merge(seub1, y = seub3, add.cell.ids = c("B01", "B03"), project = "DLIstudy")
rm(seub1, seub3);gc()
seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(seu)
seu <- ScaleData(seu, features = all.genes)
seu <- RunPCA(seu, features = VariableFeatures(object = seu))
seu <- FindNeighbors(seu, dims = 1:10)
seu <- FindClusters(seu, resolution = 0.5)
seu <- RunUMAP(seu, dims = 1:10)
DimPlot(seu, reduction = "umap", group.by = "Outcome")
DimPlot(seu, reduction = "pca", group.by = "Outcome")
rm(all.genes);gc()

```

## CODE APPLICATION

### Geneset importing

**Note:** The complete gene set list used in the paper can be downloaded from here: <https://drive.google.com/drive/folders/1yvK1RHT4fut96rFHQkHdrQGmqsxvThrw> `figure_1e_genesets.rds`

Instead of `figure_1e_genesets.rds` I used to pro inflammatory `GS.proinf.txt` geneset

```{r}
#Import the gene sets
gs1 <- read.delim("C:/Users/OZAN/Desktop/GS.proinf.txt", header = F)
gs1$V1
```

### Common Factor Analysis

Objective 1

```{r}
#Compute the CFA and correlations with mean Gene set expressions
cfaCor.test <- RunCFA_GScors(GEM = seu[["RNA"]]@data[, sample(ncol(seu[["RNA"]]@data), )], Gs = gs1$V1)
saveRDS(cfaCor.test, file = "C:/Users/OZAN/Desktop/final/cfaCor.test.rds")
#Modify Seurat object for further investigation
seu@meta.data <- cbind(seu@meta.data, cfaCor.test$Factors)
seu@meta.data %>% ggplot(aes(Factor1, Factor2, color=seu@meta.data$patient_response))+geom_point()
```

```{r}
seu@meta.data %>% ggplot(aes(Factor1, Factor3, color=seu@meta.data$patient_response))+geom_point()
```

Permutation test

```{r}
#Permutation test for observed correlation coefficients 
test.f1 <- permutationTest(Obs.cor = cfaCor.test[1]$FactCors[1], #For only first factor. Repeat for others.
                           GEM=seu[["RNA"]]@data, 
                           Factor=seu@meta.data$Factor1,
                           Gs = gs1$V1)
test.f1$p_val
hist(test.f1$Ncd)
#abline(v=cfaCor.test[1]$FactCors[1], lwd=3, col="red")

test.f2 <- permutationTest(Obs.cor = cfaCor.test[2]$FactCors[2], #For only first factor. Repeat for others.
                           GEM=seu[["RNA"]]@data, 
                           Factor=seu@meta.data$Factor2,
                           Gs = gs1$V1)
test.f2$p_val
hist(test.f2$Ncd)
#abline( v=cfaCor.test[2]$FactCors[2], lwd=3, col="blue")

test.f3 <- permutationTest(Obs.cor = cfaCor.test[3]$FactCors[3], #For only first factor. Repeat for others.
                           GEM=seu[["RNA"]]@data, 
                           Factor=seu@meta.data$Factor3,
                           Gs = gs1$V1)
test.f3$p_val
hist(test.f3$Ncd)
#abline( v=cfaCor.test[3]$FactCors[3], lwd=3, col="green")

```

### Weighted one-sided t test

Objective 2

```{r}
library(weights)
#source("https://raw.githubusercontent.com/yasinkaymaz/BIGQA-CM/main/BIGQA-WP1-functions.R")
#The link to the seurat object is here: https://drive.google.com/drive/folders/1yvK1RHT4fut96rFHQkHdrQGmqsxvThrw 
#seuObj <- readRDS("Bachireddy.all.tcells.seurat.rds")
#Figure 2B is for only Pre-DLI samples. Therefore, need to subset.
Pre_samples <- c("B1", "B3", "B5", "B7", "B9", "B17", "B19", "B21", "B23", "B25", "B27")
seu.sub <- seu@meta.data %>% filter(sample_id %in% Pre_samples)
cluster_ratioStats <- WeiTtest(Pid = seu.sub$patient_response,
                               Groups = seu.sub$Group,
                               Clusters = seu.sub$cluster_id,
                               nTry = 2000, N_r = 6, N_nr = 5,
                               GroupNames = c("R", "NR"))
#Example plot similar to Figure 2B - cluster 4
cluster_ratioStats@ratios %>% 
  filter(Cluster == 4) %>% 
  ggplot(aes(Groups, koi, fill=Groups))+geom_boxplot()+
  ylab("Proportion of cells Pre-DLI")
```

### Meta-clusters

Objective 3

```{r}
#library(Seurat)
#library(tidyverse)
library(pheatmap)
source("~/Downloads/BIGQA-CM-main-2/BIGQA-WP1-functions.R")
#The link to the seurat object is here: https://drive.google.com/drive/folders/1yvK1RHT4fut96rFHQkHdrQGmqsxvThrw 
#seuObj <- readRDS("Bachireddy.all.tcells.seurat.rds")
#Clusters to examine:
cls <- names(table(seuObj@meta.data$cluster_id))#These are all Biscuit clusters from the paper. Stores redundancy. 
# or optionally a small subset
cls <- c(1, 7, 14, 2)
#To create the clusters distance matrix, cdf, run this wrapper function:
cdm <- ClusterDistances(seu = seuObj, cls_list = cls)
#' To visually inspect the distances on a heatmap (any alternative to pheatmap can work)
pheatmap::pheatmap(cdm)
```

```{r}
#' Once Meta-clusters are defined, the `meta.data` can be modified to add a `meta_cluster` column.
#' Then, the function `WeiTtest` can be run on "All" samples. 
cluster_ratioStats <- WeiTtest(Pid = seu$patient_response,
                               Groups = seu$Group,
                               Clusters = seu$meta_cluster,
                               nTry = 2000, N_r = 6, N_nr = 5,
                               GroupNames = c("R", "NR"))
#Example plot similar to Figure 2C - meta-cluster 1 (MC1: combination of 19th and 29th clusters. Note: There is a type in the manuscript. MC1 should have 29 instead of 28. Check Sup.Fig3C)
cluster_ratioStats@ratios %>% 
  filter(Cluster == 'MC1_19_29') %>% 
  ggplot(aes(Groups, koi, fill=Groups))+geom_boxplot()+
  ylab("Proportion of cells")
```

```{python}

```
